---
title: "Final Project - Data Science for Business"
output:
  html_document:
    df_print: paged
Author: Jayanth Mantri, Kelly Wess, Aishwarya Lakshmi Krishnan, Balazs Zoletnik
always_allow_html: yes
---

# Team Information  
**Project Title: Maximise returns on Hollywood Movie Productions**    
Names: Jayanth Mantri, Kelly Wess, Aishwarya Lakshmi Krishnan, Balazs Zoletnik  
Section : AA  
Date of Submission: 14 February 2018  

#1 Purpose - Business Problem  

The film industry invests a lot of money into producing and releasing movies but the success of a movie is uncertain until it is released. Our project aims to reduce this uncertainty by predicting the probability of success of a movie based on certain factors, such as:  
1.	Genre  
2.	Budget  
3.	Runtime   
4.	Title  
5.	Release Date/Month
6.	Run time  
7.	Lead Actors  
8.	Production House  

Factors such as popularity, vote count and number of votes are listed in the database, but these cannot be used to predict the profitability of a movie (because the data is not available in time).

The objective of our project is to predict the success of a movie in terms of total revenue and profit generated as a function of each of the above factors. This analysis we believe will help produce movies that have a good chance of success in terms of its profitability ratio. 

#2 Data Collection and Summary  

This data set was found from Kaggle. It has about 85 variables for each of the 4800 movies. 

We cleaned the data as follows:
1. Removed the white spaces in "text fields"
2. Deleted entries movies without revenue/budget data
3. Creating a variable called **"ProfitFactor"**, which is a binary variable (0/1) that indicates the attractiveness of the movie based on level of profits earned
4. Created a variable called **"Proftiablity"**, which is a ratio of revenue and budget

```{r}
pacman::p_load("caret","partykit","ROCR","lift","rpart","e1071", "lubridate")

suppressWarnings(source("library.R"))
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')


datafile_name = "MovieDatabase.csv"
ProjectData <- read.csv(datafile_name)
ProjectData_INITIAL <- ProjectData

#removing entries that have budget field as empty
ProjectData_INITIAL2 <- ProjectData_INITIAL[ProjectData_INITIAL[,2]!=0,]
#removing entries that have revenue field as empty
ProjectData_INITIAL3 <- ProjectData_INITIAL2[ProjectData_INITIAL2[,3]!=0,]

#defining profitability
ProjectData_INITIAL3$Profitability <-(ProjectData_INITIAL3$revenue / ProjectData_INITIAL3$budget)

#classifying variables
ProjectData_INITIAL3$release_date <- as.Date(ProjectData_INITIAL3$release_date)
ProjectData_INITIAL3$ProfitFactor <-as.factor(ProjectData_INITIAL3$ProfitFactor)

#defining month and year of movie, from "release date"
ProjectData_INITIAL3$Month_of_release <- month(ProjectData_INITIAL3$release_date)
#Some months weren't readble from data, so we use "0"" to define those months
ProjectData_INITIAL3$Month_of_release[is.na(ProjectData_INITIAL3$Month_of_release)] <- 0 
ProjectData_INITIAL3$Year_of_release <- day(ProjectData_INITIAL3$release_date)
#Some years weren't readble from data, so we use "0"" to define those years
ProjectData_INITIAL3$Year_of_release[is.na(ProjectData_INITIAL3$Year_of_release)] <- 0 

#classifying variables
#ProjectData_INITIAL3$Month_of_release <-as.factor(ProjectData_INITIAL3$Month_of_release)
#ProjectData_INITIAL3$Year_of_release <-as.factor(ProjectData_INITIAL3$Year_of_release)


```

#Adding new variables in R

1. We created dummy variables for Keywords, Actors and Genres
2. Based on their importance and frequency, we chose to use **top 20 Actors** and **all Genres** for our modeling

```{r}
#Adding new columns with unique Keywords
library(tidyverse)
movies <- ProjectData_INITIAL3

movies_tall <- movies %>% 
  gather(keyword_col, keyword_name, starts_with("keyword")) %>%
  select(-keyword_col) %>%
  unique() %>%
  mutate(keyword_present = 1)

unique_keywords <- movies_tall %>%
  group_by(keyword_name) %>%
  summarise(keyword_occurrences = n())

random_20_keywords <- sample_n(unique_keywords, 20)$keyword_name

movies_tall_random_20_kw <- movies_tall %>% 
  mutate(keyword_name =  if_else((keyword_name %in% random_20_keywords), keyword_name, "other")) %>%
  unique()

movies_wide <- movies_tall_random_20_kw %>%
  spread(keyword_name, keyword_present, fill = 0, sep = "_")

#Adding new columns with unique Actors
movies_tall <- movies_wide %>% 
  gather(actor_col, actor_name, starts_with("Actor")) %>%
  select(-actor_col) %>%
  unique() %>%
  mutate(actor_present = 1)

unique_actors <- movies_tall %>%
  group_by(actor_name) %>%
  summarise(actor_occurrences = n())

random_1000_actors <- sample_n(unique_actors, 8973)$actor_name

movies_tall_random_1000_actors <- movies_tall %>% 
  mutate(actor_name =  if_else((actor_name %in% random_1000_actors), actor_name, "other")) %>%
  unique()

movies_wide2 <- movies_tall_random_1000_actors %>%
  spread(actor_name, actor_present, fill = 0, sep = "_")

#Adding new columns with unique Genres
movies_tall <- movies_wide2 %>% 
  gather(Genre_col, Genre_name, starts_with("Genre")) %>%
  select(-Genre_col) %>%
  unique() %>%
  mutate(Genre_present = 1)

unique_Genre <- movies_tall %>%
  group_by(Genre_name) %>%
  summarise(Genre_occurrences = n())

random_33_genres <- sample_n(unique_Genre, 20)$Genre_name

movies_tall_genres <- movies_tall %>% 
  mutate(Genre_name =  if_else((Genre_name %in% random_33_genres), Genre_name, "other")) %>%
  unique()

ProjectData_FINAL <- movies_tall_genres %>%
  spread(Genre_name, Genre_present, fill = 0, sep = "_")

```


#Create Training and Testing Data

```{r}
#creating training and testing data
set.seed(77850) #set a random number generation seed to ensure that the split is the same everytime
#install.packages("caret")
library(caret)
ProjectData_FINAL$ProfitFactor <- as.factor(ProjectData_FINAL$ProfitFactor)
inTrain <- createDataPartition(y = ProjectData_FINAL$ProfitFactor,p = 0.75, list = FALSE)
training <- ProjectData_FINAL[ inTrain,]
testing <- ProjectData_FINAL[ -inTrain,]
```


```{r}
#XG Boost
#install.packages("xgboost")
library(xgboost)
# Package loading
require(xgboost)
require(Matrix)
require(data.table)
if (!require("vcd")) install.packages("vcd")

# Preparing data for xgboost and running the model
sparse_matrix <- sparse.model.matrix(ProfitFactor ~ budget+runtime+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War , data = training)
head(sparse_matrix)
output_vector = training[,"ProfitFactor"] == 1
model_xgboost <- xgboost(data = sparse_matrix, label = output_vector
                         ,eta = 0.3, max_depth = 10, nrounds = 1500, objective ="binary:logistic", verbose = 0)

# Visualize what variables are important to predict churn
importance <- xgb.importance(feature_names = colnames(sparse_matrix), model = model_xgboost)
importance
xgb.plot.importance(importance_matrix = importance)
```



#Logistic Regression 

```{r}
pacman::p_load("caret","partykit","ROCR","lift","rpart","e1071", "lubridate")
options(warn=-1)
# GLM Logistic Regression
model_logistic<-glm(ProfitFactor~ budget+runtime+Month_of_release+Year_of_release+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War, data = training, family="binomial"(link="logit"))

summary(model_logistic)
```

```{r}
#Finding predicitons on Testing set
logistic_pred_testing<-rep("1",nrow(testing))
logistic_probabilities_testing<-predict(model_logistic,newdata=testing,type="response") #Predict probabilities
logistic_pred_testing[as.numeric(logistic_probabilities_testing)<0.755]="0" #Predict classification. Why 0.755? - that's the average probability of a movie being succesful
confusionMatrix(logistic_pred_testing,testing$ProfitFactor) #Display confusion matrix
#OC Curve
logistic_ROC_pred <- prediction(logistic_probabilities_testing, testing$ProfitFactor)
logistic_ROC_testing <- performance(logistic_ROC_pred,"tpr","fpr") #Create ROC curve data
plot(logistic_ROC_testing) #Plot ROC curve
```


```{r}
####AUC (area under curve)
auc.tmp <- performance(logistic_ROC_pred,"auc") #Create AUC data
logistic_auc_testing <- as.numeric(auc.tmp@y.values) #Calculate AUC
logistic_auc_testing #Display AUC value: 90+% - excellent, 80-90% - very good, 70-80% - good, 60-70% - so so, below 60% - not much value

#### Lift chart
plotLift(logistic_probabilities_testing, testing$ProfitFactor, cumulative = TRUE, n.buckets = 10) # Plot Lift chart

```

```{r}
#Stepwise AIC Model
library(MASS)
model_logistic_stepwiseAIC<-stepAIC(model_logistic,direction = c("both"),trace = 1) #AIC stepwise
summary(model_logistic_stepwiseAIC)
```

```{r}
par(mfrow=c(1,4))
plot(model_logistic_stepwiseAIC) #Error plots: similar nature to lm plots
par(mfrow=c(1,1))
model_logistic_FINAL<-model_logistic_stepwiseAIC #Final model
#Finding predicitons on Testing set
logistic_pred_testing<-rep("1",nrow(testing))
logistic_probabilities_testing<-predict(model_logistic_FINAL,newdata=testing,type="response") #Predict probabilities
logistic_pred_testing[as.numeric(logistic_probabilities_testing)<0.755]="0" #Predict classification. Why 0.755? - that's the average probability of a movie being succesful
confusionMatrix(logistic_pred_testing,testing$ProfitFactor) #Display confusion matrix
#OC Curve
logistic_ROC_pred <- prediction(logistic_probabilities_testing, testing$ProfitFactor)
logistic_ROC_testing <- performance(logistic_ROC_pred,"tpr","fpr") #Create ROC curve data
plot(logistic_ROC_testing) #Plot ROC curve

####AUC (area under curve)
auc.tmp <- performance(logistic_ROC_pred,"auc") #Create AUC data
logistic_auc_testing <- as.numeric(auc.tmp@y.values) #Calculate AUC
logistic_auc_testing #Display AUC value: 90+% - excellent, 80-90% - very good, 70-80% - good, 60-70% - so so, below 60% - not much value
```
```{r}
#### Lift chart
plotLift(logistic_probabilities_testing, testing$ProfitFactor, cumulative = TRUE, n.buckets = 10) # Plot Lift chart
```

# CTREE Regression
```{r}
ctree_tree<-ctree(ProfitFactor~ budget+runtime+Month_of_release+Year_of_release+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War, data = training) 

#Run ctree on training data
plot(ctree_tree, gp = gpar(fontsize = 8)) #Plotting the tree (adjust fontsize if needed)

ctree_prediction<-predict(ctree_tree,newdata=testing, type="response") #Predict classification (for confusion matrix); default with ctree

confusionMatrix(ctree_prediction,testing$ProfitFactor) #Display confusion matrix
```

```{r}
####ROC Curve
ctree_probabilities_testing <-predict(ctree_tree,newdata=testing,type = "prob") #Predict probabilities
ctree_pred_testing <- prediction(ctree_probabilities_testing[,2], testing$ProfitFactor ) #Calculate errors
ctree_ROC_testing <- performance(ctree_pred_testing,"tpr","fpr") #Create ROC curve data
plot(ctree_ROC_testing) #Plot ROC curve
```
####AUC (area under curve)
```{r}
auc.tmp <- performance(ctree_pred_testing,"auc") #Create AUC data
ctree_auc_testing <- as.numeric(auc.tmp@y.values) #Calculate AUC
ctree_auc_testing #Display AUC value: 90+% - excellent, 80-90% - very good, 70-80% - good, 60-70% - so so, below 60% - not much value

#### Lift chart
plotLift(ctree_prediction,  testing$ProfitFactor, cumulative = TRUE, n.buckets = 10) # Plot Lift chart

```
#Lasso Regression
```{r}
library(glmnet)
#convert training data to matrix format
x <- model.matrix(ProfitFactor~budget+runtime+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War,training)
#convert class to numerical variable
y <- training$ProfitFactor
#perform grid search to find optimal value of lambda
#family= binomial => logistic regression, alpha=1 => lasso
# check docs to explore other type.measure options
cv.out <- cv.glmnet(x,y,alpha=1,family="binomial",type.measure ="mse")
#plot result
plot(cv.out)
lambda_1se <-log(cv.out$lambda.1se)
```

```{r}
#get test data
x_test <- model.matrix(ProfitFactor~budget+runtime+actor_name_RobertDeNiro+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War,testing)
#predict class, type=???class???
lasso_prob <- predict(cv.out,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
lasso_predict <- rep(0,nrow(testing))
lasso_predict[lasso_prob>.755] <- 1
#confusion matrix
table(pred=lasso_predict,true=testing$ProfitFactor)
confusionMatrix(lasso_predict,testing$ProfitFactor)

```

#Linear regression to predict revenues

```{r}
#Estimating actual Revenues - linear regression GLM
suppressWarnings(source("library.R"))
pacman::p_load("caret","partykit","ROCR","lift","rpart","e1071", "lubridate")

model_logistic<-glm(revenue ~ budget+Month_of_release+Year_of_release+runtime+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War, data = training)
summary(model_logistic)
model_logistic_prediction<-predict(model_logistic, testing) 
percent.errors <- abs(testing$revenue-model_logistic_prediction)/testing$revenue*100 #calculate absolute percentage errors
mean(percent.errors) #display Mean Absolute Percentage Error (MAPE)

#Estimating actual Revenues - linear regression Stepwise AIC

model_logistic_stepwiseAIC<-stepAIC(model_logistic,direction = c("both"),trace = 1) #AIC stepwise
summary(model_logistic_stepwiseAIC)

model_logistic_prediction<-predict(model_logistic_stepwiseAIC, testing) 

percent.errors <- abs(testing$revenue-model_logistic_prediction)/testing$revenue*100 #calculate absolute percentage errors
mean(percent.errors) #display Mean Absolute Percentage Error (MAPE)

#Estimating actual Revenues - CTREE


# CTREE Regression

ctree_tree<-ctree(revenue ~budget+runtime+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War, data=training) #Run ctree on training data
plot(ctree_tree, gp = gpar(fontsize = 8)) #Plotting the tree (adjust fontsize if needed)

ctree_prediction<-predict(ctree_tree,newdata=testing, type="response") #Predict classification (for confusion matrix); default with ctree



#Lasso Regression

library(glmnet)
#convert training data to matrix format
x <- model.matrix(revenue~budget+Year_of_release+Month_of_release+runtime+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War,training)
#convert class to numerical variable
y <- training$ProfitFactor
#perform grid search to find optimal value of lambda
#family= binomial => logistic regression, alpha=1 => lasso
# check docs to explore other type.measure options
cv.out <- cv.glmnet(x,y,alpha=1,family="multinomial", type.measure ="mse")
#plot result
plot(cv.out)
lambda_1se <-log(cv.out$lambda.1se)

#get test data
x_test <- model.matrix(revenue~budget++Year_of_release+Month_of_release+runtime+actor_name_RobertDeNiro+actor_name_RobertDeNiro+actor_name_JohnnyDepp+actor_name_NicolasCage+actor_name_DenzelWashington+actor_name_ArnoldSchwarzenegger+actor_name_Himself+actor_name_TomCruise+actor_name_TomHanks+actor_name_GeorgeClooney+actor_name_MattDamon+actor_name_SylvesterStallone+actor_name_AdamSandler+actor_name_BenStiller+actor_name_JimCarrey+actor_name_HarrisonFord+actor_name_MarkWahlberg+actor_name_BenAffleck+actor_name_EddieMurphy+actor_name_JohnTravolta+actor_name_BradPitt+Genre_name_Drama+Genre_name_Comedy+Genre_name_Action+Genre_name_Romance+Genre_name_Crime+Genre_name_Adventure+Genre_name_Family+Genre_name_ScienceFiction+Genre_name_Mystery+Genre_name_Fantasy+Genre_name_Horror+Genre_name_History+Genre_name_Thriller+Genre_name_War,testing)
#predict class, type=???class???
lasso_prob <- predict(cv.out,newx = x_test,s=lambda_1se,type="response")
#translate probabilities to predictions
lasso_predict <- rep(0,nrow(testing))
lasso_predict[lasso_prob>.755] <- 1

```

