---
title: "Project Proposal - Data Science for Business"
output:
  html_document:
    df_print: paged
Author: Jayanth Mantri, Kelly Wess, Aishwarya Lakshmi Krishnan, Balazs Zoletnik
always_allow_html: yes
---
# Team Information  
**Project Title: Maximise ROI on Hollywood Movie Productions**    
Names: Jayanth Mantri, Kelly Wess, Aishwarya Lakshmi Krishnan, Balazs Zoletnik  
Section : AA  
Date of Submission: 14 February 2018  

#1 Purpose  
By doing a regresson analysis, we want to predict the best predictors for maximising returns (profits as a percentage of investment) of a movie. 

#2 Data Collection and Summary  
This data set was found from Kaggle.  
It has about 28 variables for 5043 movies, spanning across 100 years in 66 countries. There are 2399 unique director names, and thousands of actors/actresses.
```{r}
datafile_name = "tmdb_5000_movies.csv"
ProjectData <- read.csv(datafile_name)
ProjectData <- data.matrix(ProjectData) 
ProjectData_INITIAL <- ProjectData

```

#3 List of Variables


#4 Correlation Matrices

#5 Analysis Steps

**1. Data Preparation**  
a. Combine 2 csv files  
b. Clean the data - check for consistency, eliminate blanks/"NA"s  
c. Evaluate the quality of the data available  
d. Understand individual variables  


**2 Data Analysis**  
a. General Variable Summaries  
b. Generate Correlation Matrices to understand number of independent variables







```{r}

```

